\documentclass[12pt, a4paper, oneside]{report}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{graphicx}

\usepackage{perpage} %the perpage package
\MakePerPage{footnote} %the perpage package command

\usepackage{csquotes}
\usepackage[backend=biber]{biblatex}
\addbibresource{thesis.bib}

\usepackage{color}
\usepackage{listings}
%\lstset{language=Python}

\usepackage{setspace}
\usepackage{soul}

% According to the University's guide lines
\usepackage{geometry}
\geometry{margin=3cm}
%\linespread{1.5}


\definecolor{Code}{rgb}{0,0,0}
\definecolor{Decorators}{rgb}{0.5,0.5,0.5}
\definecolor{Numbers}{rgb}{0.5,0,0}
\definecolor{MatchingBrackets}{rgb}{0.25,0.5,0.5}
\definecolor{Keywords}{rgb}{0,0,1}
\definecolor{self}{rgb}{0,0,0}
\definecolor{Strings}{rgb}{0,0.63,0}
\definecolor{Comments}{rgb}{0,0.63,1}
\definecolor{Backquotes}{rgb}{0,0,0}
\definecolor{Classname}{rgb}{0,0,0}
\definecolor{FunctionName}{rgb}{0,0,0}
\definecolor{Operators}{rgb}{0,0,0}
\definecolor{Background}{rgb}{0.98,0.98,0.98}

%\lstnewenvironment{python}[1][]{
\lstset{
numbers=left,
numberstyle=\footnotesize,
numbersep=1em,
xleftmargin=1em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
frame=l,
tabsize=4,
% Basic
basicstyle=\ttfamily\small\setstretch{1},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert},
keywordstyle={\color{Keywords}\bfseries},
% additional keywords
morekeywords={[2]@invariant},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
%
}
%}{}


\usepackage{tikz}


\title{
      \textbf{Distributed Flow Control and Intelligent Data Transfer in High Performance Computing Networks}
}
\author{\large\sc{Mehdi Sadeghi}}

\date{31. March 2015}


% Defining a chapter counter
\newcounter{chapter_count}
\setcounter{chapter_count}{8}

% Moved here to change the order of package loading to prevent footnotes to bring up title page
\usepackage[colorlinks=true,
	    linkcolor=blue,
	    filecolor=magenta,
	    urlcolor=cyan,]{hyperref}

\newenvironment{dedication}
  {\clearpage           % we want a new page
   \thispagestyle{empty}% no header and footer
   \vspace*{\stretch{1}}% some space at the top 
   \itshape             % the text is in italics
   %\raggedleft          % flush to the right margin
  }
  {\par % end the paragraph
   \vspace{\stretch{3}} % space at bottom is three times that at the top
   \clearpage           % finish off the page
  }

\begin{document}
\nocite{*} % Everything in bibtex file will appear in bibliography

% Numbering according to 
% http://www.cgu.edu/PDFFiles/Office%20of%20Records%20forms/Degree%20Completion/PaginatingYourDissertation.pdf
%\pagenumbering{gobble}% Remove page numbers (and reset to 1)

\setlength{\unitlength}{1cm}

\makeatletter
\begin{titlepage}%
\rule{0mm}{1mm}
\vspace*{20mm}
%\noindent Left side \hfill right side\\
\begin{center}%
  \setlength{\unitlength}{1cm}
  \begin{picture}(12,4)
  %\thinlines\put(0,0){\framebox(12,4){}}
  \put(0,1.8){\parbox{12cm}{\centering\LARGE\bf%
    \setlength{\baselineskip}{20pt}%
    \@title}}
  \end{picture}
    \vskip 10mm%
    {\large\sc\@author}
    \vskip 10mm%
    {Supervisors:}\\
    Prof. Dr. Katharina \sc{Mehner-Heindl}\\
    Dr. Adham \sc{Hashibon}
    \vskip 30mm%
    {\Large M\,A\,S\,T\,E\,R\, T\,H\,E\,S\,I\,S}
    \vskip 20mm%
    {\large On} \vskip 1mm {Master Degree Program}%
    \vskip 4mm%
    {\large\sc{Communication and Media Engineering}}%
    \vskip 4mm%
    {\large in \\{University of Applied Sciences Offenburg}}%
    \vskip 16mm%
    {\large March 2015}%
\end{center}\par
  \vfil\null
\end{titlepage}%
\makeatother


\chapter*{Declaration of Authorship}
\noindent
I declare in lieu of an oath that the Master Thesis submitted has been produced by me without illegal help
from other persons. I state that all passages which have been taken out of publications of all means or un-published
material either whole or in part, in words or ideas, have been marked as quotations in the relevant passage. 
I also confirm that the quotes included show the extent of the original quotes and are marked
as such. I know that a false declaration will have legal consequences. \\


March 31, 2015 \\

Mehdi Sadeghi

\thispagestyle{empty} 
\pagenumbering{roman}
\clearpage

\chapter*{Copyright}
\noindent
\begin{flushleft}
\copyright 2015 Mehdi Sadeghi - sadeghi [at] mehdix [dot] org.
 
This work is licensed under a Creative Commons Attribution- ShareAlike 3.0 License.
To view a copy of this license visit:
\url{http://creativecommons.org/licenses/by-sa/3.0/legalcode}.
\end{flushleft}

\thispagestyle{empty} 

\chapter*{Abstract}
This document contains my master thesis report, including problem definition, requirements, problem analysis,
review of current state of the art, proposed solution, designed prototype, discussions and conclusion.

During this work we propose a collaborative solution to run different types of operations 
in a broker-less network without relying on a central orchestrator. 

Based on our requirements, we define and analyze a number of scenarios.
Then we design a solution to address those scenarios using a distributed workflow management approach. 
We explain how we break a complicated operation into simpler parts 
and how we manage it in a non-blocking and distributed way.
Then we show how we asynchronously launch them on the network and 
how we collect and aggregate results.

Later on we introduce an implemented prototype that demonstrates the proposed design.

\thispagestyle{empty} 
\clearpage

\begin{dedication}
To Rozita
\end{dedication}


\chapter*{Acknowledgement}
Foremost, I would like to express my gratitude to my advisor Dr. Adham Hashibon for his continuous support, creativity,
encouragement and enthusiasm during the course of this thesis.

Besides my advisor, I would like to sincerely thank my supervisor Prof. Dr. Katharina Mehner-Heindl for her insightful comments,
scientific support, and encouragement. Her support has always helped me to continue with more self-confidence.

\include{preface}

\setcounter{tocdepth}{1}
\tableofcontents
\newpage
\listoffigures
\newpage
\lstlistoflistings
\newpage

\pagenumbering{arabic}
\include{introduction} % objectives, the problem, the context

\include{requirements} % in detail
\include{literature} % making readers familiar with state of the art, where to asses?
\include{analysis} % problem analysis
\include{proposal} % explaining the proposed design/algorithm
\include{prototype} % about the implementation details and techniques
\include{discussions} % discussing possible issues and future works
\include{conclusion} % thesis conclusion and future work


\printbibliography[heading=bibintoc, title={References}]

\end{document}